---
title: "p8105_hw2_ms5965"
author: "Mehr Shafiq"
output: github_document
---

Loading libraries.

```{r setup, }
library(tidyverse)
library(readxl)
```

## Problem 1

Defining a path to the dataset.

```{r Importing data}
path_to_data = "./data/trash_wheel_data.xlsx"
```

Reading Mr Trashwheel Data.

```{r cleaning trashwheel}
trashwheel_df = 
	read_xlsx(
		path = path_to_data,
		sheet = "Mr. Trash Wheel",
		range = cell_cols("A:N")) %>% 
	janitor::clean_names() %>% 
	drop_na(dumpster) %>% 
	mutate(
		sports_balls = round(sports_balls),
		sports_balls = as.integer(sports_balls))
```
Reading and cleaning precipitation data for 2017 & 2018

```{r cleaning precipitation}
prec_17 = 
  read_excel(
    path = path_to_data,
    sheet = "2017 Precipitation",
    skip = 1) %>% 
  
  janitor::clean_names() %>% 
  drop_na(month) %>% 
  mutate(year = 2017) %>% 
  relocate (year)

prec_18 = 
  read_excel(
    path = path_to_data,
    sheet = "2018 Precipitation",
    skip = 1) %>% 
  
  janitor::clean_names() %>% 
  drop_na(month) %>% 
  mutate(year = 2018) %>% 
  relocate (year)

```

Combining 2017 & 2018 Precipitation Datasets

```{r combining precipitaion ds}
month_df = 
	tibble(
		month = 1:12,
		month_name = month.name)

prec_17_18 = bind_rows(prec_17, prec_18) 

prec_17_18 = left_join(prec_17_18, month_df, by = "month")

```

This dataset contains information from the Mr. Trashwheel trash collector in Baltimore, Maryland. As trash enters the inner harbor, the trashwheel collects that trash, and stores it in a dumpster. The dataset contains information on year, month, and trash collected, include some specific kinds of trash. There are a total of `r nrow(trashwheel_df)` observations in our final dataset. Additional data sheets include month precipitation data. The median number of sports balls found in a dumpster in 2017 was `r trashwheel_df %>% filter(year == 2017) %>% pull(sports_balls) %>% median()`. The total precipitation in 2018 was `r prec_17_18 %>% filter(year == 2018) %>% pull(total) %>% sum()` inches.


## Problem 2

Defining a path to the dataset, then cleaning it..

```{r}
path_to_NYC_data = read_csv(file = "./data/NYC_subway_data.csv") %>% 
janitor::clean_names() 
  
subset_NYC_data = path_to_NYC_data %>% 
  
  select(line, entrance_latitude, entrance_longitude, route1:route11, entry, vending, entrance_type, ada, station_name, exit_only) %>% 
  
mutate(entry = ifelse(entry == "YES", TRUE, FALSE)) 
  
```
This dataset contains information on the NYC Transit - with respect to their entrance, exit, routes, lines and ADA compliance for each subway in the city. 

The unadulterated dataset was cleaned in the following steps:
*   A subset of the dataset with only relevant variables was created.
*   The 'entry' variable which had YES & NO values was converted into a logical variables where YES -> TRUE and NO -> FALSE. 

This data has `r ncol(subset_NYC_data)` variables and `r nrow(subset_NYC_data)` observations. 

This dataset is not tidy. 

#### Problem 2: Questions

##### Part A: 

```{r}
distinct_stations = distinct(subset_NYC_data, line, station_name, .keep_all = TRUE) %>% 
  arrange(station_name, line)
```

There are `r nrow(distinct_stations)` distinct stations by name and line in NYC. 


##### Part B: 

```{r}
ADA_compliant = filter(subset_NYC_data, ada == TRUE)
```

`r nrow(ADA_compliant)` stations are ADA compliant.


##### Part C: 

```{r}
vendor_data = filter(subset_NYC_data, vending == "YES") %>% 
  count(entry)

vendor_data 
```

`r vendor_data[2,2]` out of `r sum(vendor_data$n)` stations with vending allow entrance. 


#### Tidying the dataframe

```{r}
tidy_data = mutate_at(subset_NYC_data, vars(route1:route11), as.character) %>%

pivot_longer(
    route1:route11, 
    names_to = "route_name",
    names_prefix = "route",
    values_to = "route_number") %>% 
drop_na(route_number)

```
*   The multiple route names (route1 to route11) were all converted in the same type of variable, that is, character variable.
*   The originally wide dataframe was changed into a long table where the umbrella variable "route_name" stores all the route numbers. 

This dataframe is tidy now.

##### A train

```{r}
A_train = filter(tidy_data, route_number == "A") 

A_stations = distinct(A_train, station_name, .keep_all = TRUE)

A_ADA = filter(A_stations, ada == TRUE)
```

There are `r nrow(A_stations)` distinct stations serving the A train. Of these stations, `r nrow(A_ADA)` are ADA compliant. 


## Problem 3

Cleaning the pols-month data:

```{r}
pols_month = read_csv(file = "./data/pols-month.csv") %>% 
  
  janitor::clean_names() %>% 
  
  separate(mon, into = c("year", "month", "day")) %>% 
  
  mutate(
    month = as.numeric(month),
    month = month.name[month],
    president = recode(prez_gop, `1` = "gop", `2` = "gop",`0` = "dem")) %>% 
  
  select(-day) %>%  
  select(year, month, prez_dem, -prez_gop, everything()) %>% 
  arrange(year, month)
```

This dataset contained variables related to the number of national politicians who are demogratic or republican at any given time --> these variables included year, president of that year (democratic or republican), governors, and senators. This cleaned dataset has `r nrow(pols_month)` variabels and `r ncol(pols_month)` observations. 

Cleaning the snp data:

```{r}
snp_data = read_csv(file = "./data/snp.csv") %>% 
  
  janitor::clean_names() %>% 
  
  separate(date, into = c("month", "day", "year")) %>% 
    
  mutate(
    month = as.numeric(month),
    month = month.name[month]) %>% 
    
  select(-day) %>% 
  select(year, month, everything()) %>% 
    
  arrange(year, month)
  
```

SNP - Standard & Poor's stock market index (S&P) data is used as a representative measure of stock market as a whole. This dataset had two variables: date and close (the closing values of the S&P stock index on the associated date), and `r ncol(snp_data)` observations.

Cleaning the unemployement data:

```{r}
unemployment_data = read_csv(file = "./data/unemployment.csv") %>% 
  
  janitor::clean_names()
  
tidy_unemployment = unemployment_data %>% 
  pivot_longer(
    jan:dec,
    names_to = "month",
    values_to = "unemployment"
  ) %>% 
    
  mutate(
    month = recode(month, "may" = "May"),
    month = tools::toTitleCase(month),
    month = match(month, month.abb),
    month = as.numeric(month),
    month = month.name[month]) %>% 

  select(year, month, everything()) %>% 
    
  arrange(year, month)
```

This dataset contains the percentage of employment per month in a particular year. It contains the variables: year, month and umployment (%). The cleaned version of the dataset has `r ncol(tidy_unemployment)` observations.

