---
title: "p8105_hw2_ms5965"
author: "Mehr Shafiq"
output: github_document
---

Loading libraries.

```{r setup, }
library(tidyverse)
library(readxl)
```

## Problem 1

Defining a path to the dataset.

```{r Importing data}
path_to_data = "./data/trash_wheel_data.xlsx"
```

Reading Mr Trashwheel Data.

```{r cleaning trashwheel}
trashwheel_df = 
	read_xlsx(
		path = path_to_data,
		sheet = "Mr. Trash Wheel",
		range = cell_cols("A:N")) %>% 
	janitor::clean_names() %>% 
	drop_na(dumpster) %>% 
	mutate(
		sports_balls = round(sports_balls),
		sports_balls = as.integer(sports_balls))
```
Reading and cleaning precipitation data for 2017 & 2018

```{r cleaning precipitation}
prec_17 = 
  read_excel(
    path = path_to_data,
    sheet = "2017 Precipitation",
    skip = 1) %>% 
  
  janitor::clean_names() %>% 
  drop_na(month) %>% 
  mutate(year = 2017) %>% 
  relocate (year)

prec_18 = 
  read_excel(
    path = path_to_data,
    sheet = "2018 Precipitation",
    skip = 1) %>% 
  
  janitor::clean_names() %>% 
  drop_na(month) %>% 
  mutate(year = 2018) %>% 
  relocate (year)

```

Combining 2017 & 2018 Precipitation Datasets

```{r combining precipitaion ds}
month_df = 
	tibble(
		month = 1:12,
		month_name = month.name)

prec_17_18 = bind_rows(prec_17, prec_18) 

prec_17_18 = left_join(prec_17_18, month_df, by = "month")

```

This dataset contains information from the Mr. Trashwheel trash collector in Baltimore, Maryland. As trash enters the inner harbor, the trashwheel collects that trash, and stores it in a dumpster. The dataset contains information on year, month, and trash collected, include some specific kinds of trash. There are a total of `r nrow(trashwheel_df)` observations in our final dataset. Additional data sheets include month precipitation data. The median number of sports balls found in a dumpster in 2017 was `r trashwheel_df %>% filter(year == 2017) %>% pull(sports_balls) %>% median()`. The total precipitation in 2018 was `r prec_17_18 %>% filter(year == 2018) %>% pull(total) %>% sum()` inches.


## Problem 2

Defining a path to the dataset, then cleaning it..

```{r}
path_to_NYC_data = read_csv(file = "./data/NYC_subway_data.csv") %>% 
janitor::clean_names() 
  
subset = c("line", "entrance_latitude", "entrance_longitude", "route1", "route2", "route3", "route4", "route5", "route6", "route7", "route8", "route9", "route10", "route11", "entry", "vending", "entrance_type", "ada", "station_name")

subset_NYC_data = path_to_NYC_data[subset]

```

